---
title: "Check MR1 calibration spring 2023"
author: "Baptiste schmid"
format: html
editor: visual
params:
  outdir: "C:/Users/bs/OneDrive - Vogelwarte/31403 Radartechnik/SBRS divers/Calibration_MR1/Croatia_2023"
execute:
  echo: false
---

```{r packages}
#| echo: false
#| message: false
library(tidyverse, quiet = TRUE)
library(knitr, quiet = TRUE)
# library(kableExtra))
library(dbplyr, quiet = TRUE)
library(yaml, quiet = TRUE)
library(dplyr, quiet = TRUE)
library(birdscanR, quiet = TRUE)

# Load required libraries
library(ggplot2)
library(dplyr)
library(tidyr)
library(gridExtra)
```

```{r}
mainOutputDir <- params$outdir
dir.create(mainOutputDir, showWarnings = FALSE, recursive = TRUE)
```

## Calibrate MR1

```{r}
radarID_siteName <- c("radarID", "siteName")[1]
```

Multiple MR1 radar can run in close proximity (few meters apart) to check and correct their calibration. Analyses are then conducted per `r radarID_siteName` .

If the aim is to compare the performance of a radar across time, then analyse are conducted per `r radarID_siteName` .

## Load Database Configuration

```{r setup calibration 2025}
#| eval: true
#| warning: false
# Load the YAML configuration
config_yaml <- "
databases:
  - dbname: 'HR_Palud_2023'
    startTime: '2023-03-01 14:00'
    stopTime: '2023-05-31 00:00'
    siteLocation: c(45.03970278, 13.697478)
    radarTimeZone: 'Etc/GMT0'
  
  - dbname: 'HR_Rovinj_2023'
    startTime: '2023-03-01 14:00'
    stopTime: '2023-05-31 00:00'
    siteLocation: c(45.0826367, 13.6370857)
    radarTimeZone: 'Etc/GMT0'
  
  - dbname: 'HR_Vrana_2023'
    startTime: '2023-03-01 14:00'
    stopTime: '2023-05-31 00:00'
    siteLocation: c(44.1204628, 15.5097716)
    radarTimeZone: 'Etc/GMT0'
  
  - dbname: 'HR_Zadar_2023'
    startTime: '2023-03-01 14:00'
    stopTime: '2023-05-31 00:00'
    siteLocation: c(47.494427, 15.2404117)
    radarTimeZone: 'Etc/GMT0'
"

# Parse the YAML configuration
config <- yaml.load(config_yaml)

# Convert to data frame for easier manipulation
db_config <- do.call(rbind, lapply(config$databases, data.frame))

print("Database Configuration Loaded:")
# print(db_config)
```

## Database Summary

```{r summary}
#| echo: false
# Display configuration summary
cat("Number of databases:", nrow(db_config), "\n")
cat("Date range:", unique(db_config$startTime), "to", unique(db_config$stopTime), "\n")
cat("Database names:\n")
for(i in 1:nrow(db_config)) {
  cat("-", db_config$dbname[i], "\n")
}
```

## Process Each Database

```{r set database user and pwd}
#| warning: false
dbUser =  'sa' # rstudioapi::askForPassword("Database user") # can be hard coded, e.g. 'sa'
dbPwd = 'radar' # rstudioapi::askForPassword("Database password") # can be hard oded, e.g. '123pwd'

```

```{r extract DB data with birdscanR}
#| warning: false
#| message: false

db_rds_file <- paste0("Calibration_dbExtract", paste(db_config$dbname, collapse = "_"), ".rds")
echo_rds_file <- paste0("Calibration_EchoData", paste(db_config$dbname, collapse = "_"), ".rds")
echo_rds_path <- file.path(mainOutputDir, echo_rds_file)

if(file.exists(echo_rds_path)){
  echo <- readRDS(echo_rds_path)
} else{

  # # query the databases
  for(i in 1:nrow(db_config)) { # i <- 1
    i_db <- db_config[i, ] 
    cat("Processing database:", i_db$dbname, "\n")
    cat("  Start:", i_db$startTime, "\n") 
    cat(" Stop:", i_db$stopTime, "\n") 
    
    #- extrcat data from DB
    i_dbData = extractDbData(
      dbDriverChar = "SQL Server",
      dbServer = "dbradar\\RadarData",
      dbName = i_db$dbname,
      dbUser = dbUser,
      dbPwd = dbPwd,
      saveDbToFile = FALSE,
      dbDataDir = mainOutputDir,
      radarTimeZone = i_db$radarTimeZone,
      targetTimeZone = i_db$radarTimeZone,
      #timeInterval = c(i_db$startTime, i_db$stopTime)
      listOfRfFeaturesToExtract = c(167, 168),
      siteLocation = c(47.494427, 8.716432), # i_db$siteLocation
      sunOrCivil = "civil",
      crepuscule = "nauticalSolar"
     ) 
    
  i_echo <- i_dbData$echoData %>% 
    left_join(., i_dbData$protocolData, by = 'protocolID') %>% 
    mutate(radarID = unique(i_dbData$siteData$radarID),
           transmitPower = unique(i_dbData$siteData$transmitPower),
           year = year(dateSunset)
           ) %>% 
    filter(time_stamp_targetTZ >= i_db$startTime & time_stamp_targetTZ <= i_db$stopTime) %>% 
    select(echoID, protocolID, stc_level, time_stamp_originTZ, time_stamp_targetTZ, time_string, dateSunset, year, dayOrNight, dayOrCrepOrNight, 
           fs, feature1.altitude_AGL, feature2.azimuth, feature3.speed, feature14.maxLevel, feature17.rcs, feature19.durationOfEcho, feature37.speed, 
           WFF_credibility, WFF_predicted, class, class_probability, mtr_factor_rf, #mtr_factor_sphereDiaCm, 
           siteID, pulseType, rotate, stc, threshold, radarID, transmitPower)
      
  
    if(i == 1){
      echo = i_echo
      ls_dbData <- list()
      ls_dbData[i_db$dbname] <- i_dbData
    } else {
      echo = rbind(echo, i_echo)
      ls_dbData[i_db$dbname] <- i_dbData
    }
    
    rm(list = ls(pattern = "^i_"))
    cat("\n") 
  }
  # save echo dataset to mainoutputdir
  saveRDS(echo, echo_rds_path)
  saveRDS(ls_dbData,  file.path(mainOutputDir, db_rds_file))
}
```

#### Filter Data

Fitler data for short-pulse only, and using the same threshold values. Selecting the highest threshold across DBs ensure that, given a proper radar calibration, the smallest echo disapear uniformaely across all. This is true above the STC. make sure STC is uniformally applied, or check data above the hightest STC.

Which threshold is used across the `r radarID_siteName` ?

```{r}
if(radarID_siteName == 'radarID'){
  table(echo$stc, echo$threshol, echo$pulseType, echo$radarID) 
} else {
  table(echo$stc, echo$threshol, echo$pulseType, echo$siteName) 
}
```

```{r}
# Filter the data based on the given criteria
pulseType_sel <- "S"
threshold_highest <- max(echo$threshold, na.rm = TRUE) # [dBm] selecting the highest threshold ensure that, if the radar are calibrated, the smallest echo disapear uniformaely across all. This is true above teh STC. make sure STC is uniformally applied, or check data above the hightest STC.
class_sel = 'BioClasses'
filtered_data <- echo %>%
  filter(
    pulseType == pulseType_sel,
    feature14.maxLevel >= threshold_highest,
    if (any(class_sel %in% 'BioClasses')) {
      !class %in% c('nonbio', 'precipitation')
    } else {
      class %in% class_sel
    }
  )
```

```{r get counts of echoes}
# Create table with counts per radarID for each date and day/night
date_daynight_table <- filtered_data %>%
  select(radarID, dateSunset, dayOrNight) %>%
  count(dateSunset, dayOrNight, radarID) %>%
  pivot_wider(names_from = radarID, values_from = n, values_fill = 0)

# Display date/day-night table
kable(date_daynight_table, caption = "Counts per Radar ID by Date and Day/Night")

```

#### Counts per class for the DAY with maximal number of echoes

```{r}
# Find the N (n = 1) dates (dateSunset x dayOrNight) with maximum counts
t_dayOrNight = 'day'
max_dates <- filtered_data %>%
  filter(dayOrNight == t_dayOrNight) %>% 
  select(dateSunset, dayOrNight) %>%
  count(dateSunset, dayOrNight) %>%
  arrange(desc(n)) %>%
  slice_head(n = 1)
t_date <- as.Date(max_dates$dateSunset)


# Filter echo data to only include the two dates with maximum counts
echo_filtered <- filtered_data %>%
  semi_join(max_dates, by = c("dateSunset", "dayOrNight"))

# Create the summarized and pivoted table per class and radarID using filtered data
echo_table <- echo_filtered %>%
  select(radarID, class) %>%
  count(class, radarID) %>%
  pivot_wider(names_from = radarID, values_from = n, values_fill = 0)

# Display class table as HTML table
t_captation = paste0(t_date, " ", t_dayOrNight, "time - Echo Class by Radar ID")
kable(echo_table, caption = t_captation)
```

#### Counts per class for the NIGHT with maximal number of echoes

```{r}
# Find the N (n = 1) dates (dateSunset x dayOrNight) with maximum counts
t_dayOrNight = 'night'
max_dates <- filtered_data %>%
  filter(dayOrNight == t_dayOrNight) %>% 
  select(dateSunset, dayOrNight) %>%
  count(dateSunset, dayOrNight) %>%
  arrange(desc(n)) %>%
  slice_head(n = 1)
t_date <- as.Date(max_dates$dateSunset)

# Filter echo data to only include the two dates with maximum counts
echo_filtered <- filtered_data %>%
  semi_join(max_dates, by = c("dateSunset", "dayOrNight"))

# Create the summarized and pivoted table per class and radarID using filtered data
echo_table <- echo_filtered %>%
  select(radarID, class) %>%
  count(class, radarID) %>%
  pivot_wider(names_from = radarID, values_from = n, values_fill = 0)

# Display class table as HTML table
t_captation = paste0(t_date, " ", t_dayOrNight, "time - Echo Class by Radar ID")
kable(echo_table, caption = t_captation)
```

```{r 2023-04-19}
t_date <- as.Date("2023-04-19")

# Filter echo data to only include the two dates with maximum counts
echo_filtered <- filtered_data %>%
  semi_join(max_dates, by = c("dateSunset", "dayOrNight"))

# Create the summarized and pivoted table per class and radarID using filtered data
echo_table <- echo_filtered %>%
  select(radarID, class) %>%
  count(class, radarID) %>%
  pivot_wider(names_from = radarID, values_from = n, values_fill = 0)

# Display class table as HTML table
t_captation = paste0(t_date, " ", t_dayOrNight, "time - Echo Class by Radar ID")
kable(echo_table, caption = t_captation)
```

```{r 2023-04-21}
t_date <- as.Date("2023-04-21")

# Filter echo data to only include the two dates with maximum counts
echo_filtered <- filtered_data %>%
  semi_join(max_dates, by = c("dateSunset", "dayOrNight"))

# Create the summarized and pivoted table per class and radarID using filtered data
echo_table <- echo_filtered %>%
  select(radarID, class) %>%
  count(class, radarID) %>%
  pivot_wider(names_from = radarID, values_from = n, values_fill = 0)

# Display class table as HTML table
t_captation = paste0(t_date, " ", t_dayOrNight, "time - Echo Class by Radar ID")
kable(echo_table, caption = t_captation)
```

```{r fun RCS distribution}
library(ggplot2)
library(dplyr)
library(patchwork)

plot_rcs_altitude_distributions <- function(
  data, 
  style = c("boxplot", "density", "points"),
  classes = "passerine_type",  # can be a vector of classes
  max_RCS = Inf                # new argument for maximum RCS value to include
) {
  style <- match.arg(style)

  data <- data %>%
    filter(class %in% classes,
           WFF_predicted > 8,
           WFF_predicted < 12,
           WFF_credibility > 0.4,
           dayOrNight == 'night',
           feature17.rcs <= max_RCS) %>%
    select(feature1.altitude_AGL, feature17.rcs, radarID, class, dateSunset) %>%
    mutate(radarID = as.factor(radarID),
           class = as.factor(class))
  
  class_lab <- paste0("class: ", paste(levels(droplevels(data$class)), collapse=", "))
  rcs_lab <- if(is.finite(max_RCS)) paste0("RCS ≤ ", max_RCS) else "all RCS"
  title_suffix <- paste0(class_lab, "; ", rcs_lab)
  
  # RCS plot (upper panel)
  if (style == "boxplot") {
    rcs_plot <- ggplot(data, aes(x = radarID, y = feature17.rcs, fill = radarID)) +
      geom_boxplot(alpha = 0.5, outlier.alpha = 0.2, width = 0.6) +
      labs(title = paste("RCS distribution by radarID (", title_suffix, ")", sep=""), x = "Radar ID", y = "RCS") +
      theme_minimal() +
      theme(legend.position = "none") +
      facet_wrap(~class)
  } else if (style == "density") {
    rcs_plot <- ggplot(data, aes(x = feature17.rcs, color = radarID, fill = radarID)) +
      geom_density(alpha = 0.25) +
      labs(title = paste("RCS distribution by radarID (", title_suffix, ")", sep=""), x = "RCS", y = "Density") +
      theme_minimal() +
      facet_wrap(~class)
  } else if (style == "points") {
    rcs_plot <- ggplot(data, aes(x = radarID, y = feature17.rcs, color = radarID)) +
      geom_jitter(alpha = 0.4, width = 0.2, height = 0, size = 1.1) +
      labs(title = paste("RCS distribution by radarID (", title_suffix, ")", sep=""), x = "Radar ID", y = "RCS") +
      theme_minimal() +
      theme(legend.position = "none") +
      facet_wrap(~class)
  }

  # Altitude plot (lower panel)
  if (style == "boxplot") {
    alt_plot <- ggplot(data, aes(x = radarID, y = feature1.altitude_AGL, fill = radarID)) +
      geom_boxplot(alpha = 0.5, outlier.alpha = 0.2, width = 0.6) +
      labs(title = paste("Altitude distribution by radarID (", title_suffix, ")", sep=""), x = "Radar ID", y = "Altitude AGL") +
      theme_minimal() +
      theme(legend.position = "none") +
      facet_wrap(~class)
  } else if (style == "density") {
    alt_plot <- ggplot(data, aes(x = feature1.altitude_AGL, color = radarID, fill = radarID)) +
      geom_density(alpha = 0.25) +
      labs(title = paste("Altitude distribution by radarID (", title_suffix, ")", sep=""), x = "Altitude AGL", y = "Density") +
      theme_minimal() +
      facet_wrap(~class)
  } else if (style == "points") {
    alt_plot <- ggplot(data, aes(x = radarID, y = feature1.altitude_AGL, color = radarID)) +
      geom_jitter(alpha = 0.4, width = 0.2, height = 0, size = 1.1) +
      labs(title = paste("Altitude distribution by radarID (", title_suffix, ")", sep=""), x = "Radar ID", y = "Altitude AGL") +
      theme_minimal() +
      theme(legend.position = "none") +
      facet_wrap(~class)
  }

  # Combine plots vertically
  rcs_plot / alt_plot
}
```

```{r rcs for passerine}
plot_rcs_altitude_distributions(filtered_data, style = "boxplot")
plot_rcs_altitude_distributions(filtered_data, style = "density")
```

```{r rcs for all birds}
plot_rcs_altitude_distributions(filtered_data, style = "boxplot", classes = c("passerine_type", "wader_type", "unid_bird"), max_RCS = 0.02)
plot_rcs_altitude_distributions(filtered_data, style = "density", classes = c("passerine_type", "wader_type", "unid_bird"), max_RCS = 0.02)

```

```{r}
# Provided parameters
TS <- -93 # dBm as for Pr_min 
Hmax <- 300 # m as for STCdist 
Psend <- 22 # kW as for Pt 
Again <- 20 # dBi as for G0 

# Provided function
funMinRCS <- function(TS=numeric(), Hmax=numeric(), Psend=numeric(), Again=numeric()){
  min_rcs <- (10^(TS/10) * 10^-3 * Hmax^4 * (4*pi)^3)/
    ((Psend*10^3) * (10^(Again/10))^2 * (3*10^8/(9.4*10^9))^2)   
  out <- min_rcs 
  return(out)
}

min_rcs <- funMinRCS(TS, Hmax, Psend, Again)

# Create a function for the line: y = f(x), where x is altitude (AGL)
# Reverse-engineer what the line should be for the plot (for demonstration, I’ll plot a horizontal line at min_rcs)
# If you need a more complex line, provide the explicit formula.

#library(ggplot2)
library(gridExtra)

# First plot: stc_level vs. feature1.altitude_AGL
p1 <- ggplot(df, aes(x = feature1.altitude_AGL, y = stc_level)) +
  geom_point() +
  geom_hline(yintercept = min_rcs, color = "blue", linetype = "dashed") +
  labs(title = "stc_level vs. feature1.altitude_AGL") +
  theme_bw()

# Second plot: feature14.maxLevel vs. altitude (with red for stc_level > maxLevel)
df$color_flag <- ifelse(df$stc_level > df$feature14.maxLevel, "red", "black")

p2 <- ggplot(df, aes(x = feature1.altitude_AGL, y = feature14.maxLevel)) +
  geom_point(aes(color = color_flag)) +
  scale_color_identity() +
  geom_hline(yintercept = min_rcs, color = "blue", linetype = "dashed") +
  labs(title = "feature14.maxLevel vs. altitude", y = "feature14.maxLevel") +
  theme_bw()

# Arrange the two plots
grid.arrange(p1, p2, ncol = 1)
```

```{r}
summary_table <- filtered_data %>%
  mutate(diff = stc_level - feature14.maxLevel) %>%
  group_by(radarID) %>%
  summarise(
    total_echoes = n(),
    echoes_stc_gt_feature14 = sum(stc_level > feature14.maxLevel, na.rm = TRUE),
    percent_stc_gt_feature14 = 100 * echoes_stc_gt_feature14 / total_echoes,
    q25_diff = quantile(diff[stc_level > feature14.maxLevel], 0.25, na.rm = TRUE),
    q50_diff = quantile(diff[stc_level > feature14.maxLevel], 0.5, na.rm = TRUE),
    q75_diff = quantile(diff[stc_level > feature14.maxLevel], 0.75, na.rm = TRUE)
  )
```

We here see the percentage of echoes with a max_level below the stc_level, and the quantiles of the difference between the two values.

```{r}

knitr::kable(summary_table, digits = 2, caption = "Echoes with stc_level > feature14.maxLevel per radarID")
```

```{r NOT RUN process issue dbConnect}
#| eval: false
#| echo: false

rds_name <- paste0("Calibration_EchoDataset", paste(db_config$dbname, collapse="_"), ".rds")
rds_path <- file.path(mainOutputDir, rds_name)
if(file.exists(file.path())){
  echo <- readRDS(rds_path)
} else {
  
  for(i in 1:nrow(db_config)) { # i <- 1
    db <- db_config[i, ] 
    cat("Processing database:", db$dbname, "\n")
    cat("  Start:", db$startTime, "\n") 
    cat(" Stop:", db$stopTime, "\n") 
    
    #- Connect to SQL-server
    dbServer = "dbradar\\RadarData" # "server\\instance"
    dbName = db$dbname
    dbDriverChar = "SQL Server"
    
    # For PC configuration
    # dsn = paste0(
    #       "driver=", dbDriverChar, ";server=", dbServer,
    #       ";database=", dbName,
    #       ";uid=", dbUser,
    #       ";pwd=", dbPwd
    # )
    # dbConnection = RODBC::odbcDriverConnect(dsn)
    
    con <- DBI::dbConnect(odbc::odbc(),
                          Driver = 'SQL Server',
                          Server = dbServer,
                          Database = dbName,
                          UID = dbUser,
                          PWD = dbPwd,
                          # Trusted_Connection = "Yes",
                          Port =1433
    )
    dbConnection = DBI::dbConnect(con)
    
   
    # save the connection to a table as an object
    #- Get radar and site tables from the database
    dbo_collection <- tbl(dbConnection, "collection") 
  
    dbo_rf_class <- tbl(dbConnection, "rf_classification") 
   
    dbo_rf_classprob <- tbl(dbConnection, "rf_class_probability") 
  
    dbo_rf_classdef <- tbl(dbConnection, "rfclasses") 
  
    dbo_rf_features <- tbl(dbConnection, "echo_rffeature_map") 
  
    dbo_rf_featdef <- tbl(dbConnection, "rffeatures") 
  
    dbo_protocol <- tbl(dbConnection, "protocol") 
  
    
    # Fetch the data
    echo_i <-
      dbo_collection %>% 
      full_join(., dbo_rf_class, by = c("row" = "echo"), suffix = c("", ""))  %>%
      full_join(., dbo_rf_classdef, by = c("class" = "id"), suffix = c("", ""))  %>%
      full_join(., dbo_rf_classprob, by = c("row" = "echo"), suffix = c("", "_prob"))  %>% # add class prob and MTR-factor
      full_join(., dbo_rf_features, by = c("row" = "echo"), suffix = c("", "_feature"))  %>% # add two columns: 'feature' und 'value_feature'
      filter(feature == "109" | feature == "167" | feature == "168") %>% # select a few features of interest from dbo_rf_feature$feature
      full_join(., dbo_protocol, by = "protocolID", suffix = c("", "_protocol"))  %>%
      select(my_col$colname_org) %>%
      rename_all(~my_col$colname_new) %>%
      tbl_df %>%
      spread("feature_key", "feature_value") %>% # from long to wide format of the single class probabilities
      collect() # only then the sql-query will be performed.
    echo_i <- rename("rcs_14" = "109","wbf" = "167", "wbf_cred" = "168")
    
    if(i == 1){
      echo = echo_i
    } else {
      echo = rbind(echo, echo_i)
    }
    
    cat("\n") 
  }
  saveRDS(echo, rds_path)
}

```

## Visualize

```{r scatter plot}
#| warning: false
#| echo: false

# Create a data frame with unique transmitPower values for each radarID and year combination
transmitPower_labels <- filtered_data %>%
  group_by(radarID, year) %>%
  summarize(
    transmitPower = first(transmitPower),
    # Calculate positions for upper left corner of each panel
    x_pos = min(feature14.maxLevel),  # Leftmost x value
    y_pos = max(feature1.altitude_AGL),     # Topmost y value
    .groups = "drop"
  )

# Create the plot with the transmitPower labels added in the upper left corner
passerine_plot <- ggplot(filtered_data, aes(x = feature14.maxLevel, y = feature1.altitude_AGL)) +
  # Add points
  geom_point(alpha = 0.3, color = "blue") +
  # Add transmitPower label in the upper left corner of each panel
  geom_text(
    data = transmitPower_labels,
    aes(x = x_pos, y = y_pos, label = paste0("Psent: ", transmitPower, " kW")),
    hjust = 0, vjust = 1,  # Horizontal justify left (0), vertical justify top (1)
    color = "black", size = 3, fontface = "bold",
    nudge_x = 0.5,  # Small nudge to prevent text from touching the axis
    nudge_y = -5    # Small nudge down from the top edge
  ) +
  # Facet by radarID and year (keep original faceting)
  facet_grid(year ~ radarID) +
  # Add labels and title
  labs(
    title = paste0(paste(class_sel, collapse ="-"), ", pulseType ", pulseType_sel, ", threshold ", threshold_highest),
    subtitle = "Height vs Max Level with Density Isolines",
    x = "Maximum Reflectivity Level (dBZ)",
    y = "Height (m)",
    caption = "Note: Isolines represent density quantiles (99%, 95%, 90%, 75%, 50%)"
  ) +
  # Theme adjustments
  theme_minimal() +
  theme(
    panel.grid.minor = element_blank(),
    strip.background = element_rect(fill = "lightgray"),
    strip.text = element_text(face = "bold"),
    legend.position = "bottom"
  )

# Save the plot to a file
# For a basic PNG file
file_name <- paste0( "Dbz vs height for", paste(class_sel, collapse ="-"), ".png" )
ggsave(filename = file.path(mainOutputDir, file_name), 
       plot = passerine_plot, width = 12, height = 8, dpi = 300)

# Display the plot
print(passerine_plot)




# If there are too many radarID and year combinations, consider creating a function
# to generate plots for specific combinations or using a loop
# For example:

# plot_radar_year <- function(data, radarID_val, year_val) {
#   subset_data <- data %>%
#     filter(radarID == radarID_val, year == year_val)
#   
#   ggplot(subset_data, aes(x = feature14.maxLevel, y = feature1.altitude_AGL)) +
#     geom_point(alpha = 0.3, color = "blue") +
#     geom_density_2d(aes(color = after_stat(level)), bins = 4) +
#     scale_color_viridis_c(option = "plasma", name = "Density") +
#     labs(
#       title = paste("Passerine Type Detections for Radar", radar_id_val, "in", year_val),
#       subtitle = "Height vs Max Level with Density Isolines",
#       x = "Maximum Reflectivity Level (dBZ)",
#       y = "Height (m)"
#     ) +
#     theme_minimal()
# }

# Example usage:
# plot_radar_year(filtered_data, 3220, 2023)



```

```{r}
#| warning: false
#| echo: false

# Function to extract quantiles for a given height range
extract_height_quantiles <- function(data, height = 500, spread = 25, 
                                   quantiles = c(1, 0.99, 0.95, 0.9, 0.75)) {
  
  # Define height range
  height_min <- height - spread
  height_max <- height + spread
  
  # Filter data for the specified height range
  height_filtered <- data %>%
    filter(feature1.altitude_AGL >= height_min & 
           feature1.altitude_AGL <= height_max)
  
  # Check if we have data in this height range
  if(nrow(height_filtered) == 0) {
    warning(paste("No data found in height range", height_min, "-", height_max, "m"))
    return(NULL)
  }
  
  # Calculate quantiles for each radar and year combination
  quantile_results <- height_filtered %>%
    group_by(radarID, year) %>%
    summarise(
      n_observations = n(),
      height_range = paste0(height_min, "-", height_max, "m"),
      q100 = quantile(feature14.maxLevel, probs = 1, na.rm = TRUE),
      q99 = quantile(feature14.maxLevel, probs = 0.99, na.rm = TRUE),
      q95 = quantile(feature14.maxLevel, probs = 0.95, na.rm = TRUE),
      q90 = quantile(feature14.maxLevel, probs = 0.9, na.rm = TRUE),
      q75 = quantile(feature14.maxLevel, probs = 0.75, na.rm = TRUE),
      .groups = 'drop'
    ) %>%
    # Round to 2 decimal places for cleaner display
    mutate(across(starts_with("q"), ~ round(.x, 2)))
  
  return(quantile_results)
}

# Function to create a nicely formatted table
# create_quantile_table <- function(quantile_data, title_suffix = "") {
#   
#   if(is.null(quantile_data)) {
#     return(NULL)
#   }
#   
#   # Create a formatted table using kable
#   library(knitr)
#   library(kableExtra)
#   
#   formatted_table <- quantile_data %>%
#     kable(
#       caption = paste0("Quantiles of Maximum Reflectivity Level (dBZ) by Height Range", title_suffix),
#       col.names = c("Radar ID", "Year", "N Obs", "Height Range", 
#                    "Q100 (Max)", "Q99", "Q95", "Q90", "Q75"),
#       align = c("l", "c", "c", "c", "c", "c", "c", "c", "c")
#     ) %>%
#     kable_styling(
#       bootstrap_options = c("striped", "hover", "condensed", "responsive"),
#       full_width = FALSE,
#       position = "center"
#     ) %>%
#     add_header_above(c(" " = 4, "Quantiles (dBZ)" = 5)) %>%
#     column_spec(1:2, bold = TRUE) %>%
#     column_spec(3, color = "blue") %>%
#     column_spec(5:9, background = "lightblue", alpha = 0.3)
#   
#   return(formatted_table)
# }
# Alternative simpler version without kable formatting (if kableExtra not available)
simple_quantile_table  <- function(data, height = 500, spread = 25) {

  results <- extract_height_quantiles(data, height = height, spread = spread)

  if(is.null(results)) {
    return(NULL)
  }

  cat("\nQuantiles of Maximum Reflectivity Level (dBZ)\n")
  cat("Height Range:", height - spread, "-", height + spread, "m\n")
  cat("=" , rep("=", 60), "\n", sep = "")

  print(results, n = Inf)

  #return(results)
}

# Example usage function
analyze_height_range <- function(data, height = 500, spread = 25) {
  
  cat("Analyzing height range:", height - spread, "-", height + spread, "m\n")
  cat("=================================================\n\n")
  
  # Extract quantiles
  results <- extract_height_quantiles(data, height = height, spread = spread)
  
  if(is.null(results)) {
    cat("No data available for analysis.\n")
    return(NULL)
  }
  
  # Print summary statistics
  cat("Summary:\n")
  cat("- Total radar-year combinations:", nrow(results), "\n")
  cat("- Total observations:", sum(results$n_observations), "\n")
  cat("- Radars included:", paste(unique(results$radarID), collapse = ", "), "\n")
  cat("- Years included:", paste(sort(unique(results$year)), collapse = ", "), "\n\n")
  
  # Create and display table
  # table_output <- create_quantile_table(results, 
  #                                      paste0(" (", height - spread, "-", height + spread, "m)"))
  table_output <- simple_quantile_table (data, height = height, spread = spread)
  
  # Also return raw results for further analysis
  list(
    table = table_output,
    data = results,
    summary = list(
      height_range = paste0(height - spread, "-", height + spread, "m"),
      n_combinations = nrow(results),
      total_obs = sum(results$n_observations),
      radars = unique(results$radarID),
      years = sort(unique(results$year))
    )
  )
}

# Quick usage examples:
# 
# # Analyze 500m ± 25m height range
# results_500m <- analyze_height_range(filtered_data, height = 500, spread = 25)
# print(results_500m$table)
# 
# # Analyze 1000m ± 50m height range  
# results_1000m <- analyze_height_range(filtered_data, height = 1000, spread = 50)
# print(results_1000m$table)
# 
# # Access raw data for further analysis
# raw_data <- results_500m$data
# View(raw_data)
```

#### Max level at 500 m agl

```{r}
# # Analyze 500m ± 25m height range
results_500m <- analyze_height_range(filtered_data, height = 500, spread = 25)
print(results_500m$table)
```

```{r plot height distribution}

# Assuming your data is loaded as 'filtered_data'
# If you need to load it, uncomment and modify the following line:
# filtered_data <- read.csv("your_data_file.csv")

# Data preparation
radar_data <- filtered_data %>%
  filter(!is.na(feature1.altitude_AGL) & !is.na(class) & !is.na(radarID)) %>%
  mutate(
    radarID = factor(radarID),
    class = factor(class),
    dayOrNight = factor(dayOrNight)
  )

# Get unique radar IDs and classes for consistent plotting
radar_ids <- sort(unique(radar_data$radarID))
classes <- sort(unique(radar_data$class))

print(paste("Number of radar IDs:", length(radar_ids)))
print(paste("Classes:", paste(classes, collapse = ", ")))
print(paste("Day/Night categories:", paste(unique(radar_data$dayOrNight), collapse = ", ")))

# Function to create frequency plots
create_frequency_plot <- function(data, day_night_filter = NULL) {
  
  plot_data <- data
  title_suffix <- ""
  
  if (!is.null(day_night_filter)) {
    plot_data <- data %>% filter(dayOrNight == day_night_filter)
    title_suffix <- paste(" -", str_to_title(day_night_filter))
  }
  
  p <- ggplot(plot_data, aes(x = feature1.altitude_AGL)) +
    geom_histogram(bins = 50, alpha = 0.7, color = "black", size = 0.3) +
    facet_grid(class ~ radarID, scales = "free_y") +
    labs(
      title = paste("Frequency Distribution of Echo Heights", title_suffix),
      x = "Altitude AGL (m)",
      y = "Frequency"
    ) +
    theme_minimal() +
    theme(
      strip.text = element_text(size = 8),
      axis.text = element_text(size = 7),
      plot.title = element_text(size = 12, hjust = 0.5)
    )
  
  return(p)
}

# Function to create density plots
create_density_plot <- function(data, day_night_filter = NULL) {
  
  plot_data <- data
  title_suffix <- ""
  
  if (!is.null(day_night_filter)) {
    plot_data <- data %>% filter(dayOrNight == day_night_filter)
    title_suffix <- paste(" -", str_to_title(day_night_filter))
  }
  
  p <- ggplot(plot_data, aes(x = feature1.altitude_AGL)) +
    geom_density(alpha = 0.7, fill = "skyblue", color = "darkblue") +
    facet_grid(class ~ radarID, scales = "free_y") +
    labs(
      title = paste("Density Distribution of Echo Heights", title_suffix),
      x = "Altitude AGL (m)",
      y = "Density"
    ) +
    theme_minimal() +
    theme(
      strip.text = element_text(size = 8),
      axis.text = element_text(size = 7),
      plot.title = element_text(size = 12, hjust = 0.5)
    )
  
  return(p)
}

# Function to create relative density plot per radarID (horizontal)
create_relative_density_plot <- function(data, day_night_filter = NULL) {
  
  plot_data <- data
  title_suffix <- ""
  
  if (!is.null(day_night_filter)) {
    plot_data <- data %>% filter(dayOrNight == day_night_filter)
    title_suffix <- paste(" -", str_to_title(day_night_filter))
  }
  
  # Calculate relative densities
  density_data <- plot_data %>%
    group_by(radarID, class) %>%
    do({
      if(nrow(.) > 1) {
        d <- density(.$feature1.altitude_AGL, na.rm = TRUE)
        data.frame(
          altitude = d$x,
          density = d$y / max(d$y)  # Normalize to relative density
        )
      } else {
        data.frame(altitude = numeric(0), density = numeric(0))
      }
    }) %>%
    ungroup()
  
  p <- ggplot(density_data, aes(y = altitude, x = density)) +
    geom_line(color = "darkred", size = 1) +
    geom_area(alpha = 0.3, fill = "red") +
    facet_grid(class ~ radarID, scales = "free") +
    labs(
      title = paste("Relative Density of Echo Heights per Radar ID", title_suffix),
      y = "Altitude AGL (m)",
      x = "Relative Density"
    ) +
    theme_minimal() +
    theme(
      strip.text = element_text(size = 8),
      axis.text = element_text(size = 7),
      plot.title = element_text(size = 12, hjust = 0.5)
    )
  
  return(p)
}

# Generate plots for all data
print("Creating frequency plot for all data...")
freq_all <- create_frequency_plot(radar_data)

print("Creating density plot for all data...")
density_all <- create_density_plot(radar_data)

print("Creating relative density plot for all data...")
rel_density_all <- create_relative_density_plot(radar_data)

# Display plots
print(freq_all)
print(density_all)
print(rel_density_all)

# Generate plots filtered by day/night
if("day" %in% radar_data$dayOrNight) {
  print("Creating plots for day data...")
  freq_day <- create_frequency_plot(radar_data, "day")
  density_day <- create_density_plot(radar_data, "day")
  rel_density_day <- create_relative_density_plot(radar_data, "day")
  
  print(freq_day)
  print(density_day)
  print(rel_density_day)
}

if("night" %in% radar_data$dayOrNight) {
  print("Creating plots for night data...")
  freq_night <- create_frequency_plot(radar_data, "night")
  density_night <- create_density_plot(radar_data, "night")
  rel_density_night <- create_relative_density_plot(radar_data, "night")
  
  print(freq_night)
  print(density_night)
  print(rel_density_night)
}

# Summary statistics
print("Summary statistics by radarID and class:")
summary_stats <- radar_data %>%
  group_by(radarID, class, dayOrNight) %>%
  summarise(
    count = n(),
    mean_altitude = mean(feature1.altitude_AGL, na.rm = TRUE),
    median_altitude = median(feature1.altitude_AGL, na.rm = TRUE),
    sd_altitude = sd(feature1.altitude_AGL, na.rm = TRUE),
    min_altitude = min(feature1.altitude_AGL, na.rm = TRUE),
    max_altitude = max(feature1.altitude_AGL, na.rm = TRUE),
    .groups = 'drop'
  )

print(summary_stats)

# Optional: Save plots to files
# Uncomment the following lines if you want to save the plots
ggsave(file.path(mainOutputDir, "frequency_all.png"), freq_all, width = 12, height = 8, dpi = 300)
ggsave(file.path(mainOutputDir, "density_all.png"), density_all, width = 12, height = 8, dpi = 300)
ggsave(file.path(mainOutputDir, "relative_density_all.png"), rel_density_all, width = 12, height = 8, dpi = 300)

```
